---
title: 01-Haiti wvs2016 data cleaning
subtitle: relig-civ-eng-Haiti paper
author: "Raph M."
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    css: lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
    smooth_scroll: true
    number_sections: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

# Load packages

```{r load-pkgs, echo =TRUE, include=TRUE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages (library(foreign))
suppressPackageStartupMessages (library(data.table))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages (library(broom))
suppressPackageStartupMessages (library(skimr))
suppressPackageStartupMessages (library(janitor))
# install.packages("devtools")
# devtools::install_github("thomasp85/patchwork")
suppressPackageStartupMessages (library(patchwork))
```
<br>

# Call the data file

```{r 01-data, echo = TRUE, include = TRUE, warning = FALSE}
setwd("/cloud/project/religion-civic-Haiti/code/raw-code")
# 
bagay <- read.dta("WV6_Haiti_2016.dta")

## Convert it into data.table for more efficient data analyses
bagay <- data.table(bagay)

# skim(bagay)

dim(bagay)
# head(bagay[, 25:35])


#  define a simpler function for tabulations that shows all NAs
ftab <- function(x){
  a <- table(x, useNA = "always")
  return (a)
}
```
<br>

# Get a sense of missingness in this data set

```{r missing, echo = TRUE, include=TRUE, warning=FALSE, results='markup', message=FALSE}
numo <- nrow(bagay)
ncomplete = sum(complete.cases(bagay))
ncomplete/numo
# 1   ==> which means 100% complete 
#  this shows that every row is complete and that there are no missing values. but that's grossly misleading. 

# install.packages("naniar")
suppressPackageStartupMessages (library(naniar))
## visualize missingness
# vis_miss(bagay[1:100])  # uncomment and play with the range as needed

# This shows clearly that: The missing values here are coded not as NA but as several different options depending on the variables.

# So we will have to deal with missingness one variable at a time as we encounter them.  

```
<br>

# Recode variables 
    
    
## Secular Association Participation  
<br>
I detail out each specific secular group as a sub-dummy before coding it like Sarkissian did ==> as an overall dummy: 1 = respondent belongs to any of the secular groups below; 0 = they don't belong.

I will then use these sub-dummies to code the overall dummy for secular organizations later.   
<br>

### Part 1: secular sub-dummies 
<br>

<!-- When you wantt your code to shoiw but not the results ==> use this chunk header -->

```{r secular-sub, echo =TRUE, include=TRUE, warning=FALSE, results='hide'}

# %%%%%%%%%%%%%%%% PART 1: SUB-DUMMIES %%%%%%%%%%%%%%%%

# V26. Sport or recreational organization
bagay$sport_member <- NA
bagay$sport_member[bagay$V26 == "Not a member"] <- 0
bagay$sport_member[bagay$V26 == "Inactive member" | bagay$V26 =="Active member"] <- 1
ftab(bagay$V26)
ftab(bagay$sport_member)
#  ok

# V27. Art, music or educational organization
bagay$art_member <- NA
bagay$art_member[bagay$V27 == "Not a member"] <- 0
bagay$art_member[bagay$V27 == "Inactive member" | bagay$V27 =="Active member"] <- 1
ftab(bagay$V27)
ftab(bagay$art_member)
#  ok

# V28. Labor Union
bagay$union_member <- NA
bagay$union_member[bagay$V28 == "Not a member"] <- 0
bagay$union_member[bagay$V28 == "Inactive member" | bagay$V28 =="Active member"] <- 1
ftab(bagay$V28)
ftab(bagay$union_member)
#  ok

# V29. Political party
bagay$pol_party_member <- NA
bagay$pol_party_member[bagay$V29 == "Not a member"] <- 0
bagay$pol_party_member[bagay$V29 == "Inactive member" | bagay$V29 =="Active member"] <- 1
ftab(bagay$V29)
ftab(bagay$pol_party_member)
#  ok

# V30. Environmental organization
bagay$environ_member <- NA
bagay$environ_member[bagay$V30 == "Not a member"] <- 0
bagay$environ_member[bagay$V30 == "Inactive member" | bagay$V30 =="Active member"] <- 1
ftab(bagay$V30)
ftab(bagay$environ_member)
#  ok

# V31. Professional association
bagay$pro_member <- NA
bagay$pro_member[bagay$V31 == "Not a member"] <- 0
bagay$pro_member[bagay$V31 == "Inactive member" | bagay$V31 =="Active member"] <- 1
ftab(bagay$V31)
ftab(bagay$pro_member)
# ok

# V32. Humanitarian or charitable organization
bagay$charity_member <- NA
bagay$charity_member[bagay$V32 == "Not a member"] <- 0
bagay$charity_member[bagay$V32 == "Inactive member" | bagay$V32 =="Active member"] <- 1
ftab(bagay$V32)
ftab(bagay$charity_member)
#  ok

# V33. Consumer organization
bagay$consumer_member <- NA
bagay$consumer_member[bagay$V33 == "Not a member"] <- 0
bagay$consumer_member[bagay$V33 == "Inactive member" | bagay$V33 =="Active member"] <- 1
ftab(bagay$V33)
ftab(bagay$consumer_member)

# V34. Self-help group, mutual aid group
bagay$self_help_member <- NA
bagay$self_help_member[bagay$V34 == "Not a member"] <- 0
bagay$self_help_member[bagay$V34 == "Inactive member" | bagay$V34 =="Active member"] <- 1
ftab(bagay$V34)
ftab(bagay$self_help_member)
#  ok

# V35. Other organization
bagay$other_org_member <- NA  
bagay$other_org_member[bagay$V35 == "Not a member"] <- 0
bagay$other_org_member[bagay$V35 == "Inactive member" | bagay$V35 =="Active member"] <- 1
ftab(bagay$V35)
ftab(bagay$other_org_member)

```
<br>

### Part 2: overall secular dummy 

```{r secular-overall, warning=FALSE, echo = TRUE, results='hide'}
names(bagay)
# idx <- bagay %>% select(relig_member, sport_member, art_member, union_member, pol_party_member, environ_member, pro_member,     
#                         charity_member, consumer_member, self_help_member, other_org_member)
# names(idx)

dim(bagay)
# 
# With a for loop I would need this:
# I need to fill a new column in the df as follows:
# for every row in df {
# do this: 
# -  look at the value in each of these variables {11 variables}
# - if the value in *any one of these variables* is 1 assign the value 1 to the new variable
# - if the value in *all of all of these variables is 0* assign the value 0 to the new variable 
# -- if the value in *all of all of these variables is NA* assign the value NA to the new variable 
# 		
# }

bagay$secular <- NA
bagay$secular[ bagay$sport_member == 0
              & bagay$art_member == 0 & bagay$union_member == 0
              & bagay$pol_party_member == 0 & bagay$environ_member == 0 
              & bagay$pro_member ==0 & bagay$charity_member == 0 
              & bagay$consumer_member == 0 & bagay$self_help_member == 0 
              & bagay$other_org_member == 0]  <- 0 
  
bagay$secular[ bagay$sport_member == 1 
              |  bagay$art_member == 1 |  bagay$union_member == 1
              |  bagay$pol_party_member == 1 |  bagay$environ_member == 1
              |  bagay$pro_member ==1  |  bagay$charity_member == 1 
              |  bagay$consumer_member == 1 |  bagay$self_help_member == 1 
              |  bagay$other_org_member == 1]  <- 1

ftab(bagay$secular)

head(bagay[1:4, 431:440])
str(bagay$secular)

# bagay$tmp <- as.factor(bagay$secular)
# ftab(bagay$tmp)
# bagay <- bagay %>% select(-tmp)
# names(bagay)

#  prelim. exploratory plot 0.1
# ggplot(bagay, aes(x=secular, na.rm = TRUE)) + 
#   geom_bar(fill=c("navy", "brown")) + 
#   labs(x= "membership in secular organization") +
#   theme_bw()

# ok
```
<br>

#### Save data at this stage so that we do not need to read it from raw data each time

```{r saving1, warning=FALSE, echo = TRUE, results='hide'}
save(bagay, file="WVS-Haiti2016.csv")

#  we will keep on writing over this file to update it at various stages of the data clenaing process.
```    
<br>  

## Signed a petition, joined in boycotts, peaceful demonstrations, strikes or other protests  
<br>
  
The 2016 WVS-Haiti contains five related questions asking participants whether or not they have signed petitions, joined in boycotts, attended peaceful demonstrations or joined strikes in the past year. These can be examined separately, each as a dummy variable (Sarkissian did in her analysis -- but she only had three such questions for the countries she picked). Or they can be combined in any manner that seems reasonable ==> for example:   

<br>

+ signed a petition or joined in boycotts
+ joined in peaceful demonstrations or strikes 
+ other protests
`This "other" category could always include things the authorities consider to be riots or unpeaceful protests`. 

<br>

`Sarkissian's prose`: "While these activities are not often used to measure participation in more established democracies, they may help to capture the potential for political engagement in the presence of more competitive politics. In the case of joining boycotts, this might include boycotting elections in which some parties are not allowed to compete, as has been common insome Middle Eastern and African countries."

<br>

I agree. These activities may reflect political engagement based on party-politics (find documentation from the New York Times that opposition candidates hire hoodlums to boycott elections or terrorize the population on election day...), socio-economic strife, etc.

<br>
```{r petition, warning=FALSE, echo = TRUE, results='hide'}
ftab(bagay$V87)
# ggplot(bagay, aes(V85)) + geom_bar()


# V85. Signing a petition
bagay$petition <- NA
bagay$petition[bagay$V85 == "Might do" | bagay$V85 =="Would never do"] <- 0
bagay$petition[bagay$V85 == "Have done"] <- 1

ftab(bagay$petition)
ftab(bagay$V85)

# V86. Joining in boycotts
bagay$boycotts <- NA
bagay$boycotts[bagay$V86 == "Might do" | bagay$V86 =="Would never do"] <- 0
bagay$boycotts[bagay$V86 == "Have done"] <- 1
ftab(bagay$boycotts)
ftab(bagay$V86)

# V87. Attending peaceful demonstrations
bagay$peaceful_march <- NA
bagay$peaceful_march[bagay$V87 == "Might do" | bagay$V87 =="Would never do"] <- 0
bagay$peaceful_march[bagay$V87 == "Have done"] <- 1
ftab(bagay$peaceful_march)
ftab(bagay$V87)

# V88. Joining strikes
bagay$strikes <- NA
bagay$strikes[bagay$V88 == "Might do" | bagay$V88 =="Would never do"] <- 0
bagay$strikes[bagay$V88 == "Have done"] <- 1
ftab(bagay$strikes)
ftab(bagay$V88)

# V89. Any other act of protest?
bagay$other_protest <- NA
bagay$other_protest[bagay$V89 == "Might do" | bagay$V89 =="Would never do"] <- 0
bagay$other_protest[bagay$V89 == "Have done"] <- 1
ftab(bagay$other_protest)
ftab(bagay$V89)
#  good!
```
  
<br>

## Voting Habits  
<br>

**QUESTION**: _"When elections take place, do you vote always, usually or never? Please tell me separately for each of the following levels"_

        1. Let's code it as follows:

             - "always" or "usually" as 1; 
             - anything else as 0. 
             -  No answer is NA

        2. Be sure to code both variables: *national* and *local political* participation  

<br>
```{r voting, warning=FALSE, echo = TRUE, results='hide'}

# V226. Local level 
bagay$voted_local <- NA
bagay$voted_local[bagay$V226 == "Never"] <- 0
bagay$voted_local[bagay$V226 == "Always" | bagay$V226 =="Usually"] <- 1
ftab(bagay$voted_local)
ftab(bagay$V226)

# V227. National level
bagay$voted_natl <- NA
bagay$voted_natl[bagay$V227 == "Never"] <- 0
bagay$voted_natl[bagay$V227 == "Always" | bagay$V227 =="Usually"] <- 1
ftab(bagay$voted_natl)
ftab(bagay$V227)
```

<br>

## Frequency of Religious Service Attendance

<br>

`V145. Apart from weddings and funerals, about how often do you attend religious services these days?`

        1 = once a week or more than once a week,  
        0 = once a month, only on special holy days, once a year, less often, never, practically never.   

<br>
```{r church-attendance, warning=FALSE, echo = TRUE, results='hide'}
#
ftab(bagay$V145)

# V145 freq_relig_services 

bagay$freq_relig_services  <- NA

bagay$freq_relig_services[  bagay$V145 == "Once a month" 
                  | bagay$V145 == "Only on special holy days" 
                  | bagay$V145 == "Once a year" 
                  | bagay$V145 == "Less often"
                  | bagay$V145 == "Never, practically never"] <- 0

bagay$freq_relig_services[bagay$V145 == "Once a week" | bagay$V145 =="More than once a week"] <- 1
# 
ftab(bagay$freq_relig_services)
ftab(bagay$V145)
# ok. 

```
<br>

## Religious Organization Membership

`“Now I am going to read off a list of voluntary organizations. For each organization,` 
`could you tell me whether you are an active member, an inactive member or not a member of that type of organization?”`  

          1 = a member (active or inactive); 
          0 = not a member of that type of organization.

<br>
```{r relig-member, warning=FALSE, echo = TRUE, results='hide'}
# V25. Church or religious organization
ftab(bagay$V25)
str(bagay$V25)
# 
bagay$relig_member <- NA  

bagay$relig_member[bagay$V25 == "Not a member"] <- 0

bagay$relig_member[bagay$V25 == "Inactive member" 
                   | bagay$V25 =="Active member"] <- 1
ftab(bagay$V25)
ftab(bagay$relig_member)

#  prelim. exploratory plot 0.2
# ggplot(bagay, aes(x=relig_member, na.rm = TRUE)) + 
#   geom_bar(fill=c("navy", "orange3")) + 
#   labs(x= "membership in a religious organization") +
#   theme_bw()
#  ok
```

<br>

## Frequency of Prayer

`Question & Raw categories`:  

`V146. Apart from weddings and funerals, about how often do you pray? (Code one answer):__`

        1. Several times a day 
        2. Once a day
        3. Several times each week
        4. Only when attending religious services
        5. Only on special holy days
        6. Once a year
        7. Less often
        8. Never, practically never

```{r prayer, echo=TRUE, warning=FALSE, results='hide'}
# Recode:
# 1 = once a day or Several times a day 
# 0 = Several times each week, Only when attending religious services,Only on special holy days 

# V146 freq_prayer 

ftab(bagay$V146)
# 
bagay$freq_prayer  <- NA

bagay$freq_prayer[  bagay$V146 == "Several times each week" 
                  | bagay$V146 == "Only on special holy days" 
                  | bagay$V146 == "Only when attending religious services" 
                  | bagay$V146 == "Only on special holy days"
                  | bagay$V146 == "Once a year"
                  | bagay$V146 == "Less often than once a year"
                  | bagay$V146 == "Never, practically never"] <- 0

bagay$freq_prayer[bagay$V146 == "Once a day" 
                          | bagay$V146 =="Several times a day"] <- 1
# 
ftab(bagay$freq_prayer)
ftab(bagay$V146)
# ok. 

```
<br>

## Social Trust

`“Generally speaking, would you say that most people can be trusted (1)` 
`or that you need to be very careful in dealing with people (0)?”`  

<br>

`V24.`   


<br>
```{r trust, echo=TRUE, warning=FALSE, results='hide'}

# V24. social trust 

ftab(bagay$V24)
# 
bagay$soc_trust  <- NA

bagay$soc_trust[  bagay$V24 == "Need to be very careful"] <- 0

bagay$soc_trust[bagay$V24 == "Most people can be trusted"] <- 1
# 
ftab(bagay$soc_trust)
ftab(bagay$V24)
# ok. 

```
<br>

## Tolerance  

 `“Here is a list of qualities that children can be encouraged to learn at home.` 
 `Which, if any, do you consider to be especially important?”`  
 
         1 = mentioned “tolerance and respect for people”;  
         0 = did not mention.  

`V16:  Tolerance and respect for other people `  

<br>
```{r tolerance, echo=TRUE, warning=FALSE, results='hide'}

# V16. tolerance 

ftab(bagay$V16)
# 
bagay$tolerance  <- NA

bagay$tolerance[bagay$V16 == "Not mentioned"] <- 0

bagay$tolerance[bagay$V16 == "Mentioned"] <- 1
# 
ftab(bagay$tolerance)
ftab(bagay$V16)
# ok. 

```

<br>

## Female Gender:   

_Sex by observation_   

          1 = female, 
          0 = male        

`V240. (respondent’s sex by observation):`

        1 Male =======> becomes 0
        2 Female =====> becomes 1

<br>
```{r female, echo=TRUE, warning=FALSE, results='hide'}

# V240. female 

ftab(bagay$V240)
# 
bagay$female  <- NA

bagay$female[bagay$V240 == "Male"] <- 0

bagay$female[bagay$V240 == "Female"] <- 1
# 
ftab(bagay$female)
ftab(bagay$V240)
# ok. 

```

<br>

## Age (in years)  

`V242. This means you are ____ years old (write in age in two digits).`


```{r age, echo=TRUE, warning=FALSE, results='hide'}

# V242. age 
# 
str(bagay$V242)
# integer  -- makes sense!
# continuous discrete. 

# Missingness ?
sum(is.na(bagay$V242))
#  0 --- great!

bagay$age  <- bagay$V242

summary(bagay$age)
summary(bagay$V242)

# ok. 

```
<br>

## Income  

            (preferred measure of SES)

`V239. On this card is an income scale on which 1 indicates the lowest income group`
`and 10 the highest income group in your country. We would like to know in what `
`group your household is. Please, specify the appropriate number, counting all `
`wages, salaries, pensions and other incomes that come in. (Code one number):`   

<br>
```{r v239, eval = FALSE, echo = TRUE}

# Lowest group <===========================================> Highest group 
#     1 	  2	   3	 4  	 5  	6   	7   	8   	 9  	10 
```
<br>

```{r income, echo=TRUE, warning=FALSE, results='hide'}

# V239. Income

ftab(bagay$V239)
# 
bagay$income_decile  <- NA
bagay$income_decile[bagay$V239 == "Lower step"] <- 1
bagay$income_decile[bagay$V239 == "second step"] <- 2
bagay$income_decile[bagay$V239 == "Third step"] <- 3
bagay$income_decile[bagay$V239 == "Fourth step"] <- 4
bagay$income_decile[bagay$V239 == "Fifth step"] <- 5
bagay$income_decile[bagay$V239 == "Sixth step"] <- 6
bagay$income_decile[bagay$V239 == "Seventh step"] <- 7
bagay$income_decile[bagay$V239 == "Eigth step"] <- 8
bagay$income_decile[bagay$V239 == "Nineth step"] <- 9
bagay$income_decile[bagay$V239 == "Tenth step"] <- 10

# 
str(bagay$income_decile)
ftab(bagay$income_decile)
ftab(bagay$V239)
# ok. 

```

<br>

## Education    

`From page 20 of documentation`: 

`V258. Weight variable (Provide a 4-digit weight variable to correct your sample `
`to reflect national distributions of key variables. If no weighting is necessary,`
`simply code each case as “1.” It is especially important to correct for education. `
`For example, if your sample contains 10 percent more university-educated respondents`
`as there are in the adult population, members of this group should be downweighted`
`by 10 percent, giving them a weight of .90).`  


        “What is the highest educational level that you have attained?    

          1 = low;    
          2 = medium;   
          3 = high.  
 
 
**It’s very unclear how Sarkissian decided to code low/medium/high education here. She did not say how she coded it. Seems arbitrary. I can do my own.**    

<br>

`V248. What is the highest educational level that you have attained?`   

**NOTE**: if respondent indicates to be a student, code highest level s/he expects to complete) 

<br>

          1. No formal education
          2. Incomplete primary school
          3. Complete primary school
          4. Incomplete secondary school: technical/vocational type
          5. Complete secondary school: technical/vocational type
          6. Incomplete secondary: university-preparatory type
          7. Complete secondary: university-preparatory type
          8. Some university-level education, without degree
          9. University-level education, with degree


<br>
```{r educ, echo=TRUE, warning=FALSE, results='hide'}

# V248. Education

ftab(bagay$V248)
# 
# 1 = low;   ===========> No formal education | Incomplete primary school | Complete primary school 

# 2 = medium; ==========>  Incomplete secondary school: technical/vocational type | Complete secondary school: technical/vocational type | Incomplete secondary: university-preparatory type

# 3 = high =============> Complete secondary: university-preparatory type | Some university-level education, without degree | University-level education, with degree

# 
bagay$educ <- NA
  
bagay$educ[ bagay$V248 == "No formal education"
               | bagay$V248 == "Incomplete primary school" 
               | bagay$V248 == "Complete primary school"] <- 1 

bagay$educ[ bagay$V248 == "Incomplete secondary school: technical/ vocational type" 
               | bagay$V248 == "Complete secondary school: technical/ vocational type" 
               | bagay$V248 == "Incomplete secondary school: university-preparatory type"] <- 2

bagay$educ[ bagay$V248 == "Complete secondary school: university-preparatory type" 
               | bagay$V248 == "Some university-level education, without degree" 
               | bagay$V248 == "University - level education, with degree"] <- 3
# 
ftab(bagay$educ)
ftab(bagay$V248)
 #  1    2    3   <NA> 
 # 659 1048  288    1 

# ok. 

```
<br>

## Subjective Social Class
`NOTE==> (Sarkissian only used this because it was difficult t obtain income data)`.  

I am recoding it here for exploratory purposes.

`V238. People sometimes describe themselves as belonging to the working class,`
`the middle class, or the upper or lower class. Would you describe yourself as`
`belonging to the (read out and code one answer):` 

            1  Upper class
            2  Upper middle class
            3  Lower middle class
            4  Working class
            5  Lower class

I am reversing the emphatic order of these levels in my own recode (e.g. upper class now will be assigned "5" as opposed to "1")

<br>
```{r class, echo=TRUE, warning=FALSE, results='hide'}

# V238. social class

ftab(bagay$V238)
# 
bagay$social_class  <- NA
bagay$social_class[bagay$V238 == "Upper class"] <- 5
bagay$social_class[bagay$V238 == "Upper middle class"] <- 4
bagay$social_class[bagay$V238 == "Lower middle class"] <- 3
bagay$social_class[bagay$V238 == "Working class"] <- 2
bagay$social_class[bagay$V238 == "Lower class"] <- 1

# 
str(bagay$social_class)
ftab(bagay$social_class)
ftab(bagay$V238)
# ok. 

```
<br>

## Lives with parents
`V250. Do you live with your parents? (Code one answer):`  

      1   Yes
      2   No

<br>
```{r lives-with-parents, echo=TRUE, warning=FALSE, results='hide'}

# V250. Lives with parents  

ftab(bagay$V250)
# 
bagay$lives_with_parents  <- NA

bagay$lives_with_parents[bagay$V250 == "No"] <- 0

bagay$lives_with_parents[bagay$V250 == "Yes"] <- 1

# 
str(bagay$lives_with_parents)
ftab(bagay$lives_with_parents)
ftab(bagay$V250)
# ok. 

```
<br>

## Population size  

`:: (NOT IN SARKISSIAN”S)::`    

`V253. (Code size of town):`  

            - 1 Under 2,000
            - 2 2,000 - 5,000
            - 3 5 - 10,000
            - 4 10 - 20,000
            - 5 20 - 50,000
            - 6 50 - 100,000
            - 7 100 - 500,000
            - 8 500,000 and more

This is already in the correct emphatic order.

<br>
```{r pop_size, echo=TRUE, warning=FALSE, results='hide'}

# V253. Population size

ftab(bagay$V253)
# 
bagay$pop_size  <- NA
bagay$pop_size[bagay$V253 == "Under 2,000"] <- 1
bagay$pop_size[bagay$V253 == "2,000-5,000"] <- 2
bagay$pop_size[bagay$V253 == "5,000-10,000"] <- 3
bagay$pop_size[bagay$V253 == "10,000-20,000"] <- 4
bagay$pop_size[bagay$V253 == "20,000-50,000"] <- 5
bagay$pop_size[bagay$V253 == "50,000-100,000"] <- 6
bagay$pop_size[bagay$V253 == "100,000-500,000"] <- 7
bagay$pop_size[bagay$V253 == "500,000 and more"] <- 8

# 
str(bagay$pop_size)
ftab(bagay$pop_size)
ftab(bagay$V253)
# ok. 

```
<br>

## Ethnicity 
`::(NOT IN SARKISSIAN”S)::`

`V254. (Code ethnic group by observation, modify for your own society):`

          - 1 Black
          - 2 White Haitian
          - 3 Mulatto
          - 4 Asian
          - 5 Other (write in): _____________________________

I am only recoding this for exploratory purposes and to get an idea of how representative the sample is.    

<br> 

According to the CIA World FactBook, Haiti's population is:   

        - 95%  black 
        - 5%   mixed & white 
        
        
<br>
```{r ethnic, echo=TRUE, warning=FALSE, results='hide'}

# V254. Ethnicity  


#  MY RECODE:
# - 1 Black
# - 2 White Haitian / mulatto
# - 3 other


ftab(bagay$V254)
# 
bagay$ethnicity  <- NA   

bagay$ethnicity[bagay$V254 == "Black-Other / Black"] <- 1

bagay$ethnicity[bagay$V254 == "HT: Mixed Race (Mullato) Haitian" | bagay$V254 == "HT: White Haitian"] <- 2

bagay$ethnicity[bagay$V254 == "Other" | bagay$V254 == "HT: Asian/Middle Eastern"] <- 3


# 
str(bagay$ethnicity)
ftab(bagay$ethnicity)
ftab(bagay$V254)
# ok. 

```
<br>

In terms of ethnic groups, the sample is representative:   

        - 1932/1996 ====> 96% Black   
        - 60/1996 ======> 3% white/Mulatto   

`The CIA estimates are from 2018 and these figures in the smaple are from 2016.`
`It should not make a big difference, but the estimates are close enough`

<br>


#### Save the tidy data again in the correct folder:
<br>
```{r save-data, warning=FALSE, include = TRUE, echo = TRUE}

# save a copy here in `raw-code` : 
save(bagay, file = "tidy-WVS-Haiti2016.csv")


# change directory to save another copy in `tidy-data`  

setwd("/cloud/project/religion-civic-Haiti/data/tidy-data")

save(bagay, file = "tidy-WVS-Haiti2016.csv")


#  return to the original directory
setwd("/cloud/project/religion-civic-Haiti/code/raw-code")


#  for reproducibility's sake, let's record the session info:
sessionInfo()
```    
<br>   



































